---
title: "contraception.csv"
format:
  html:
    theme: 
      - lumen
    css: ["../assets/codebooks.css"]
    code-copy: true
    code-fold: true
    highlight-style: zenburn
    df-print: paged
    linkcolor: '#ff851b'
date: "07-28-2022"
bibliography: '../assets/data.bib'
csl: '../assets/apa-single-spaced.csl'
---


```{r}
#| echo: false
source("../assets/codebook-setup.R")
```

---

The data in [contraception.csv](https://github.com/zief0002/epsy-8264/raw/master/data/contraception.csv) were collected from several sources (e.g., World Bank). The variables are:

- `country`: Country name
- `region`: Region of the world
- `contraceptive`: Percentage of women who are practicing, or whose sexual partners are practicing, any form of contraception. It is usually measured for women ages 15&ndash;49 who are married or in union.
- `educ_female`: Average number of years of formal education (schooling) for females
- `gni`: Categorical measure of the economy indicating if the country has a low or high gross national income

<br />


### Preview

```{r}
#| echo: true

# Import data
contraception = readr::read_csv(file = "https://raw.githubusercontent.com/zief0002/pensive-giraffe/main/data/contraception.csv")

# View data
contraception
```

<br />


### References

<div id="refs" class="references csl-bib-body hanging-indent" role="doc-bibliography">

<div class="csl-entry" role="doc-biblioentry">Roser, M. (2017). [Fertility rate](https://ourworldindata.org/fertility-rate#empowerment-of-women). *Our world in data*.</div>

<div class="csl-entry" role="doc-biblioentry">UNICEF. (2016). [State of the world's children 2016](https://www.unicef.org/publications/files/UNICEF_SOWC_2016.pdf). *United Nations Population Division's World Contraceptive Use, household surveys including Demographic and Health Surveys and Multiple Indicator Cluster Surveys.*</div>

<div class="csl-entry" role="doc-biblioentry">World Bank (2019). [World Bank open data](https://data.worldbank.org/).</div>

<div>
