---
title: "credit.csv"
format:
  html:
    theme: 
      - lumen
    css: ["../assets/codebooks.css"]
    code-copy: true
    code-fold: true
    highlight-style: zenburn
    df-print: paged
    linkcolor: '#ff851b'
date: "07-28-2022"
bibliography: '../assets/data.bib'
csl: '../assets/apa-single-spaced.csl'
---


```{r}
#| echo: false
source("../assets/codebook-setup.R")
```

---

The data in [credit.csv](https://raw.githubusercontent.com/zief0002/pensive-giraffe/main/data/credit.csv) contains simulated credit card data for 400 individuals. These data are from @James:2013. The variables in the data set are:

- `balance`: Customer's average credit card balance (in dollars)
- `income`: Customer's reported income (in \$10,000 dollars)
- `limit`: Credit limit issued to customer
- `rating`: Customer's credit rating; higher values indicate a better credit rating
- `cards`: Number of credit cards the customer has
- `age`: Customer's age
- `education`: Number of years of education

<br />


### Preview

```{r}
#| echo: true

# Import data
credit = readr::read_csv(file = "https://raw.githubusercontent.com/zief0002/pensive-giraffe/main/data/credit.csv")

# View data
credit
```

<br />


### References
