---
title: "equal-education-opportunity.csv"
format:
  html:
    theme: 
      - lumen
    css: ["../assets/codebooks.css"]
    code-copy: true
    code-fold: true
    highlight-style: zenburn
    df-print: paged
    linkcolor: '#ff851b'
date: "09-29-2022"
bibliography: '../assets/data.bib'
csl: '../assets/apa-single-spaced.csl'
---


```{r}
#| echo: false
source("../assets/codebook-setup.R")
```

---

In 1964, the US Congress passed the Civil Rights Act and also ordered a survey of school districts to evaluate the availability of equal educational opportunity in public education. The results of this survey were reported on in @Coleman:1966 and @Mosteller:1972. The data in [equal-education-opportunity.csv](https://github.com/zief0002/epsy-8264/raw/master/data/equal-education-opportunity.csv) are a subset of the data collected. These data, provided by @Chatterjee:2012, constitute a random sample of 70 schools from the original data. The attributes, which have all been mean-centered and standardized, include:

- `achievement`: Measurement indicating the student achievement level
- `faculty`: Measurement indicating the faculty's credentials
- `peer`: Measurement indicating the influence of peer groups in the school
- `school`: Measurement indicating the school facilities (e.g., building, teaching materials)

<br />


### Preview

```{r}
#| echo: true

# Import data
eeo = readr::read_csv(file = "https://raw.githubusercontent.com/zief0002/pensive-giraffe/main/data/equal-education-opportunity.csv")

# View data
eeo
```

<br />


### References


