---
title: "evaluations.csv"
format:
  html:
    theme: 
      - lumen
    css: ["../assets/codebooks.css"]
    code-copy: true
    code-fold: true
    highlight-style: zenburn
    df-print: paged
    linkcolor: '#ff851b'
date: "07-28-2022"
bibliography: '../assets/data.bib'
csl: '../assets/apa-single-spaced.csl'
---


```{r}
#| echo: false
source("../assets/codebook-setup.R")
```

---

The data in [evaluations.csv](https://raw.githubusercontent.com/zief0002/pensive-giraffe/main/data/evaluations.csv) come from @Hamermesh:2005 and were made available by @Gelman:2007. This data were collected from student evaluations of instructors' beauty and teaching quality for several courses at the University of Texas. The teaching evaluations were conducted at the end of the semester, and the beauty judgments were made later, by six students who had not attended the classes and were not aware of the course evaluations. The variables are:

- `prof_id`: Professor ID number
- `avg_eval`: Average course rating
- `num_courses`: Number of courses for which the professor has evaluations
- `num_students`: Number of students enrolled in the professor's courses
- `perc_evaluating`: Average percentage of enrolled students who completed an evaluation
- `beauty`: Measure of the professor's beauty composed  of the average score on six standardized beauty ratings
- `tenured`: Is the professor tenured? (0 = non-tenured; 1 = tenured)
- `native_english`: Is the professor a native English speaker? (0 = non-native English speaker; 1 = native English speaker)
- `age`: Professor's age (in years)
- `female`: Is the professor female? (0 = not female; 1 = female)

<br />


### Preview

```{r}
#| echo: true

# Import data
evaluations = readr::read_csv(file = "https://raw.githubusercontent.com/zief0002/pensive-giraffe/main/data/evaluations.csv")

# View data
evaluations
```


<br />

### References
