---
title: "keith2.csv"
format:
  html:
    theme: 
      - lumen
    css: ["../assets/codebooks.css"]
    code-copy: true
    code-fold: true
    highlight-style: zenburn
    df-print: paged
    linkcolor: '#ff851b'
date: "08-02-2023"
bibliography: '../assets/data.bib'
csl: '../assets/apa-single-spaced.csl'
---


```{r}
#| echo: false
source("../assets/codebook-setup.R")
```

---

The data, stored in [keith2.csv](https://raw.githubusercontent.com/zief0002/redesigned-adventure/main/data/keith2.csv) includes three attributes on $n = 32$ 8th-grade students. These data come from @Keith:2015. The attributes are:

- `gpa`: Overall Grade-point average (GPA) in all subjects (on a standard 100-point scale)
- `homework`: Average time spent on homework per week across all subjects (in hours)
- `parent_ed`: Education-level (in years of schooling) for the parent with the highest level of education


<br />


### Preview

```{r}
#| echo: true

# Import data
keith2 = readr::read_csv(file = "https://raw.githubusercontent.com/zief0002/redesigned-adventure/main/data/keith2.csv")

# View data
keith2
```

<br />


### References




