---
title: "path-model-achievement.csv"
format:
  html:
    theme: 
      - lumen
    css: ["../assets/codebooks.css"]
    code-copy: true
    code-fold: true
    highlight-style: zenburn
    df-print: paged
    linkcolor: '#ff851b'
date: "08-02-2023"
bibliography: '../assets/data.bib'
csl: '../assets/apa-single-spaced.csl'
---


```{r}
#| echo: false
source("../assets/codebook-setup.R")
```

---

The data, stored in [path-model-achievement.csv](https://raw.githubusercontent.com/zief0002/redesigned-adventure/main/data/path-model-achievement.csv) includes five standardized attributes collected from $n = 1000$ students. These data come from @Keith:2015. The attributes are:

- `achieve`: Standardized measure of achievement
- `ability`: Standardized measure of student ability level
- `motivation`: Standardized measure of motivation
- `coursework`: Standardized measure of previous coursework
- `fam_back`: Standardized measure of family background

<br />


### Preview

```{r}
#| echo: true

# Import data
path_model = readr::read_csv(file = "https://raw.githubusercontent.com/zief0002/redesigned-adventure/main/data/path-model-achievement.csv")

# View data
path_model
```

<br />


### References




