---
title: "slid.csv"
format:
  html:
    theme: 
      - lumen
    css: ["../assets/codebooks.css"]
    code-copy: true
    code-fold: true
    highlight-style: zenburn
    df-print: paged
    linkcolor: '#ff851b'
date: "07-28-2022"
bibliography: '../assets/data.bib'
csl: '../assets/apa-single-spaced.csl'
---


```{r}
#| echo: false
source("../assets/codebook-setup.R")
```

---

The data in [slid.csv](https://raw.githubusercontent.com/zief0002/pensive-giraffe/main/data/slid.csv) includes a subset of data collected during the 1994 wave of Statistics Canada's *Survey of Labour and Income Dynamics* (SLID). These data constitute employed citizens living in Ontario, Canada between the ages of 16 and 65. These data are taken from the public-use dataset made available by Statistics Canada, and prepared by the Institute for Social Research, York University. They were made available in @fox:2022. The variables in the data set are:

- `wages`: Composite hourly wage rate based on all the participant's jobs
- `age`: Age of the participant (in years)
- `education`: Number of years of schooling
- `male`: A dummy-coded predictor for sex (0=Non-male; 1=Male)

<br />


### Preview

```{r}
#| echo: true

# Import data
slid = readr::read_csv(file = "https://raw.githubusercontent.com/zief0002/pensive-giraffe/main/data/slid.csv")

# View data
slid
```

<br />


### References
