---
title: "stack-1979.csv"
format:
  html:
    theme: 
      - lumen
    css: ["../assets/codebooks.css"]
    code-copy: true
    code-fold: true
    highlight-style: zenburn
    df-print: paged
    linkcolor: '#ff851b'
date: "07-28-2022"
bibliography: '../assets/data.bib'
csl: '../assets/apa-single-spaced.csl'
---


```{r}
#| echo: false
source("../assets/codebook-setup.R")
```

---

@Stack:1979 studied predictors of income inequality in a paper published in the *American Sociological Review*. He posited that more political participation and a strong Socialist party in a country would be associated with less income inequality. To control for variation in economic development, he included a measure of energy consumption, which had been identified as a reasonable economic proxy in previous studies. Stack's data, stored in [stack-1979.csv](https://raw.githubusercontent.com/zief0002/pensive-giraffe/main/data/stack-1979.csv), include four attributes measured on 18 countries. The attributes are:

- `country`: Country name
- `inequality`: Ratio of the share of income received by the most wealthy population quintile (richest 20%) to the share received by the poorest 40% of the population; Higher values indicate more income inequality
- `turnout`: Proportion of the adult population voting in the most recent national election prior to 1972
- `energy`: Energy consumption per capita (expressed in million metric tons of coal equivalents; higher values indicate more economic development
- `socialist`: Annual average proportion of seats held by socialist parties in the national legislature, over the first twenty postwar years

<br />


### Data Preview

```{r}
#| echo: true

# Import data
stack = readr::read_csv(file = "https://raw.githubusercontent.com/zief0002/pensive-giraffe/main/data/stack-1979.csv")

# View data
stack
```

<br />


### References
