---
title: "states-2019.csv"
format:
  html:
    theme: 
      - lumen
    css: ["../assets/codebooks.css"]
    code-copy: true
    code-fold: true
    highlight-style: zenburn
    df-print: paged
    linkcolor: '#ff851b'
date: "07-28-2022"
bibliography: '../assets/data.bib'
csl: '../assets/apa-single-spaced.csl'
---


```{r}
#| echo: false
source("../assets/codebook-setup.R")
```

---

The data in [states-2019.csv](https://raw.githubusercontent.com/zief0002/pensive-giraffe/main/data/states-2019.csv) include statistics collected from Wikipedia, the 2019 American Community Survey, and the National Centers for Environmental Information. The attributes in the data are:

- `state`: State/territory name
- `life_expectancy`: Life expectancy (in years)
- `population`: Population estimate (in millions)
- `income`: Per capita income (in thousands of dollars)
- `illiteracy`: Illiteracy rate (in percent)
- `murder`: Murder and non-negligent manslaughter rate (per 100,000 population)
- `hs_grad`: Percentage of high school graduates
- `frost`: Mean number of days with minimum temperature below freezing
- `area`: Land area (in square miles)

<br />


### Preview

```{r}
#| echo: true

# Import data
usa = readr::read_csv(file = "https://raw.githubusercontent.com/zief0002/pensive-giraffe/main/data/states-2019.csv")

# View data
usa
```

<br />


### References




